---
title: "R Notebook"
output:
  html_document: default
  html_notebook: default
---

```{r setup, include=FALSE, error=TRUE, message=FALSE}
# Set knitr options for knitting code into the report:
# - Don't print out code (echo)
# - Save results so that code blocks aren't re-run unless code changes (cache),
# _or_ a relevant earlier code block changed (autodep), but don't re-run if the
# only thing that changed was the comments (cache.comments)
# - Don't clutter R output with messages or warnings (message, warning)
  # This _will_ leave error messages showing up in the knitted report

knitr::opts_chunk$set(echo = TRUE, 
                      cache = TRUE,
                      autodep = TRUE,
                      cache.comments = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      error = TRUE,
                      comment = NA)
```

## Discrete Distributions
* Binomial
* Poisson

## Binomial distribution
Some examples where the results of an experiment can be modeled by a bino-
mial random variable are:

* A drug has probability 0.90 of curing a disease. It is administered
to 100 patients, where the outcome for each patient is either cured
or not cured. If X is the number of patients cured, then X is a
binomial random variable with parameters (100, 0.90).

* The National Institute of Mental Health estimates that there is a
20% chance that an adult American suffers from a psychiatric dis-
order. Fifty adult Americans are randomly selected. If we let X
represent the number who have a psychiatric disorder, then X takes
on values according to the binomial distribution with parameters
(50, 0.20).

* A manufacturer of computer chips finds that on the average 5%
are defective. To monitor the manufacturing process, they take a
random sample of size 75. If the sample contains more than five
defective chips, then the process is stopped. The binomial distri-
bution with parameters (75, 0.05) can be used to model the random
variable X, where X represents the number of defective chips.


## Example 2.1
Suppose there is a 20% chance that an adult American suffers from a psychi-
atric disorder. We randomly sample 25 adult Americans. If we let X represent
the number of people who have a psychiatric disorder, then X is a binomial
random variable with parameters . We are interested in the proba-
bility that at most 3 of the selected people have such a disorder. We can use
the MATLAB Statistics Toolbox function binocdf to determine , as
follows:

```{r}
library(RcppOctave)
.CallOctave("version")

binocdf <- OctaveFunction("
function [b] = binocdf_()
  pkg load statistics

  b = binocdf(3, 25, 0.2);
end
")

```


```{r}
prob <- binocdf()

prob
```


We could also sum up the individual values of the probability mass function
from X = 0 to X = 3:
```{r}
binopdf <- OctaveFunction("
function [b] = binopdf_()
  pkg load statistics

  b = sum(binopdf(0:3, 25, 0.2));
end
")

prob2 <- binopdf()
prob2
```

Both of these commands return a probability of 0.234. We now show how to
generate the binomial distributions shown in Figure 2.3.

```{r}
# Get the values for the domain, x
x <- 0:6

# Get the values of the probability mass function.
# First for n = 6, p = 0.3:
pdf1 <- .O$binopdf(x, 6, 0.3)

# Now for n = 6, p = 0.7
pdf2 <- .O$binopdf(x, 6, 0.7)
```

```{r}
pdf1
pdf2
```

```{r}
library(ggplot2)
library(gridExtra)

df <- data.frame(x, pdf1, pdf2)

p1 <- ggplot(df, aes(x, pdf1)) +
  geom_bar(stat = "identity") +
  ggtitle("n = 6, p = 0.3") + 
  theme(plot.title = element_text(hjust=0.5)) +
  labs(x = "X")

p2 <- ggplot(df, aes(x, pdf2)) +
  geom_bar(stat = "identity") +
  ggtitle("n = 6, p = 0.7") +
  theme(plot.title = element_text(hjust=0.5))+ 
  labs(x = "X")

gridExtra::grid.arrange(p1, p2, ncol =2)
```

## Poisson
The Poisson distribution can be used in many applications. Examples of sit-
uations where a discrete random variable might follow a Poisson distribution
are:

* the number of typographical errors on a page,
* the number of vacancies in a company during a month, or
* the number of defects in a length of wire.


## Example 2.2
In preparing this text, we executed the spell check command, and the editor
reviewed the manuscript for typographical errors. In spite of this, some mis-
takes might be present. Assume that the number of typographical errors per
page follows the Poisson distribution with parameter $\alpha = 0.25$. We calculate the probability that a page will have at least two errors as follows:

```{r}
poisscdf <- OctaveFunction("
function [x] = poisscdf_(X, lambda)
  pkg load statistics

  x = poisscdf(X, lambda);
end
")

prob <- 1 - poisscdf(1, 0.25)
prob
```

WHich is calculated:

$$
P(X >= 2) = 1 - \{ P(X=0) + P(X=1) \} \\
= 1 - e^{-0.25} - e^{-0.25} \times 0.25 \approx 0.0265
$$

##  Example 2.3
Suppose that accidents at a certain intersection occur in a manner that satis-
fies the conditions for a Poisson process with a rate of 2 per week ($\alpha = 2$). What is the probability that at most 3 accidents will occur during the
next 2 weeks? Using Equation 2.26, we have:

As before, we can use the poisscdf function with parameter given by $\lambda t = 2 \times 2$:

```{r}
prob <- poisscdf(3, 2*2)
prob
```

## Uniform
Perhaps one of the most important distributions is the uniform distribution
for continuous random variables. One reason is that the uniform (0, 1) distri-
bution is used as the basis for simulating most random variables as we dis-
cuss in Chapter 4.

## Example 2.4
In this example, we illustrate the uniform probability density function over
the interval (0, 10), along with the corresponding cumulative distribution
function. The MATLAB Statistics Toolbox functions unifpdf and unifcdf
are used to get the desired functions over the interval.

```{r}
# First, get the domain over which we will evaluate the function
x <- seq(-1, 11, 0.1)

# implement the functions in Octave
unifpdf <- OctaveFunction("
function [p] = unifpdf_(X, a, b)
  pkg load statistics
  p = unifpdf(X, a, b);
end
")

unifcdf <- OctaveFunction("
function [p] = unifcdf_(X, a, b)
  pkg load statistics
  p = unifcdf(X, a, b);
end
")

# Get the probability density function values at x
pdf <- unifpdf(x, 0, 10)

# Now get the cdf
cdf <- unifcdf(x, 0, 10)

```


```{r fig.asp=1}
par(mfrow=c(1,2))
plot(x, pdf, type = "l", main = "PDF", ylim = c(0, 0.2), ylab = "f(X)")
plot(x, cdf, type = "l", main = "CDF", ylim = c(0, 1), ylab = "F(X)")
```

**FIGURE 2.4** On the left is a plot of the probability density function for the uniform (0, 10). Note that the height of the curve is given by . The corresponding cumulative distribution function is shown on the right.


## Normal
A well known distribution in statistics and engineering is the normal distri-
bution. Also called the Gaussian distribution, it has a continuous probability
density function given by
  <equation>

Some special properties of the normal distribution are given here.

* The value of the probability density function approaches zero as x
approaches positive and negative infinity.
* The probability density function is centered at the mean $\mu$ , and
the maximum value of the function occurs at $x = \mu$.
* The probability density function for the normal distribution is sym-
metric about the mean $\mu$


## Example 2.5
Similar to the uniform distribution, the functions `normpdf` and `normcdf` are
available in the MATLAB Statistics Toolbox for calculating the probability
density function and cumulative distribution function for the normal. There
is another special function called `normspec` that determines the probability
that a random variable X assumes a value between two limits, where X is nor-
mally distributed with mean and standard deviation This function also
plots the normal density, where the area between the specified limits is
shaded. The syntax is shown below.

```{r}
mu <- 5
sigma = 2
# set up the upper and lower limits
specs <- c(2, 8)

# the function normspec is not implemented in Octave yet.

# normspec <- OctaveFunction("
# function [p] = normspec_(specs, mu, sigma)
#   pkg load statistics
#   normspec(specs, mu, sigma);
# end
# ")

prob <- normspec(specs, mu, sigma)
prob
```

```{r}
normal = function(mean, sd, lb,ub){
  x <- seq(-4, 4, length=100) * sd + mean
  hx <- dnorm(x, mean, sd)
 
  plot(x, hx, type="n", ylab="", 
       axes = FALSE,
       main="Normal Distribution")
 
  i <- x >= lb & x <= ub
  lines(x, hx)
  polygon(c(lb, x[i], ub), c(0, hx[i], 0), col="red")
 
  area <- pnorm(ub, mean, sd) - pnorm(lb, mean, sd)
  result <- paste("P(",lb,"< IQ <",ub,") =",
                  signif(area, digits=3))
  mtext(result, 3)
  axis(1, at=seq(mean-4*sd, mean+4*sd, sd), pos=0) 
}

normal(5, 2, 2, 8)
```


## Exponential
The exponential distribution can be used to model the amount of time until a
specific event occurs or to model the time between independent events. Some
examples where an exponential distribution could be used as the model are:

* the time until the computer locks up,
* the time between arrivals of telephone calls, or
* the time until a part fails. 

## Example 2.6
The time between arrivals of vehicles at an intersection follows an exponen-
tial distribution with a mean of 12 seconds. What is the probability that the
time between arrivals is 10 seconds or less? We are given the average interarrival time, so $\lambda = 1/12$. 

We can calculate this using the Octave function `expcdf`.

```{r}
csexpoc <- OctaveFunction("
function [p] = csexpoc_(x, lambda)
  pkg load statistics
  p = expcdf(x, lambda);
end
")

prob <- csexpoc(10, 12)
prob
```

Manually, the calculation is:

$$
P(X<= 10) = 1 - e^{-(1/12)10} \approx 0.57
$$


```{r fig.asp=0.8}
gampdf <- OctaveFunction("
function [p] = gampdf_(x, t, lambdainv)
  pkg load statistics
  p = gampdf(x, t, lambdainv);
end
")

x <- seq(0, 3, 0.1)
y1 <- gampdf(x, 1, 1/1)
y2 <- gampdf(x, 2, 1/2)
y3 <- gampdf(x, 3, 1/3)

plot(x, y1, type = "l", col = "red", 
     main = "Gamma Distribution", xlab = "X", ylab = "f(x)")
lines(x, y2, col = "green")
lines(x, y3, col = "blue")
```
FIGURE 8. We show three examples of the gamma probability density function. We see that when $\lambda = t = 1$, we have the same probability density function as the exponential with parameter $\lambda = 1$

## Weibull
The Weibull distribution has many applications in engineering. In particular,
it is used in reliability analysis. It can be used to model the distribution of the amount of time it takes for objects to fail. For the special case where 
$\nu = 0$ and $\beta = 1$, the Weibull reduces to the exponential with $\lambda = 1 / \alpha$.

## Example 2.8 
Suppose the time to failure of piston rings for stream-driven compressors can
be modeled by the Weibull distribution with a location parameter of zero, $\beta = 1/3$, and $\alpha = 500$. We can find the mean time to failure using the expected value of a Weibull random variable, as follows:

```{r}
mttf <- 1/.O$wblcdf(1, 500^(-1/3), 3)
mttf
```


```{r}
# wblcdf(x, scale, shape)
prob <- .O$wblcdf(2000, 500, 1/3)
prob
```

```{r}

```

**NOTE**
In Martinez book, $\nu$ is the `location` parameter, $\alpha$ is the `scale`, and the `shape` is $\beta$.

### Weibull CDF
From the Octave manual in 2005:

**`wblcdf`(x, alpha, sigma)** \    

         Compute the cumulative distribution function (CDF) at X of the
         Weibull distribution with shape parameter ALPHA and scale
         parameter SIGMA, which is
             1 - exp(-(x/alpha)^sigma)
             for X >= 0.
       
   
   
```{r}
x <- 2000
alpha <- 500
beta <- sigma <- 1/3

1 - exp(-(x/alpha)^sigma)
```     

```{r}
.O$wblcdf(x, alpha, sigma)
```

### Weibull random matrix
`wblrnd` (ALPHA, SIGMA, R, C) \     

     Return an R by C matrix of random samples from the Weibull
     distribution with parameters ALPHA and SIGMA which must be scalar
     or of size R by C.

     If R and C are omitted, the size of the result matrix is the
     common size of ALPHA and SIGMA.
     
     
```{r}
.O$wblrnd(alpha, sigma, 4, 4)
```

### Weibull quantile

`wblinv` (X, LAMBDA, ALPHA) \   

     Compute the quantile (the inverse of the CDF) at X of the Weibull
     distribution with shape parameter ALPHA and scale parameter SIGMA.

     
```{r}
.O$wblinv(x, alpha, sigma)
```
Is this correct?

### Weibull PDF
`wblpdf` (X, ALPHA, SIGMA) \   

     Compute the probability density function (PDF) at X of the Weibull
     distribution with shape parameter ALPHA and scale parameter SIGMA
     which is given by
             alpha * sigma^(-alpha) * x^(alpha-1) * exp(-(x/sigma)^alpha)
     for X > 0.

```{r}
.O$wblpdf(x, alpha, sigma)
```

