---
title: "R Notebook"
output:
  html_document: default
  html_notebook: default
---

```{r setup, include=FALSE, error=TRUE, message=FALSE}
# Set knitr options for knitting code into the report:
# - Don't print out code (echo)
# - Save results so that code blocks aren't re-run unless code changes (cache),
# _or_ a relevant earlier code block changed (autodep), but don't re-run if the
# only thing that changed was the comments (cache.comments)
# - Don't clutter R output with messages or warnings (message, warning)
  # This _will_ leave error messages showing up in the knitted report

knitr::opts_chunk$set(echo = TRUE, 
                      cache = TRUE,
                      autodep = TRUE,
                      cache.comments = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      error = TRUE,
                      comment = NA)
```

Source: http://www.stat.wisc.edu/~bates/JSM2001.pdf


## Example 4.31. Weibull distribution
The Weibull probability plot is not available directly in R. However, the plot can be created using the formula ln[− ln(1 − p)] for the 5th, 15th, . . . , and 95th percentiles as given in text. The sequence 0.05, 0.15, . . . , 0.95 is generated with seq(0.05, 0.95, 0.1).

```{r}
library(Devore7)

data(xmp04.31)
with(xmp04.31, 
     plot(log(-log(1 - seq(0.05, 0.95, 0.1))), log(lifetime)))
```

## An R sesssion on the Weibull example
```{r}
library(Devore7)

data(xmp04.31)
str(xmp04.31)
```

```{r}
# Reasonable starting estimates are shape = 1, scale = 1000
# Do a simple evaluation at this set of parameters￿
sum(dweibull(xmp04.31$lifetime, shape=1, scale=1000, log=TRUE))
```

```{r}
# Optimization functions minimize so use negative log-likelihood
llfunc <- function(x) { # express as a function
 -sum(dweibull(xmp04.31$lifetime, shape=x[1], scale=x[2], log=TRUE))
}

# apply non-linear minimization
mle <- nlm(llfunc, c(shape = 1, scale = 1000), hessian = TRUE)

```

## Results of the Weibull Example

```{r}
str(mle)
```


```{r}
solve(mle$hessian) # approximate variance-covariance matrix￿İ
```

We see that the maximum of the log-likelihood is -77.1 achieved
at $\alpha = 2.15$ and $\beta = 1289.34$. The approximate standard errors of the estimates are $0.5444 = \sqrt 0.29596$ and $199.6 = \sqrt 38935.82$. We can use the standard errors to determine a grid of $(\alpha, \beta)$ values for contouring the log-likelihood function.

## Plotting the desnsity at the estimates

```{r}
plot(function(x) dweibull(x, shape = 2.15, scale = 1289.34), 0, 3000,
col = "red", xlab = "lifetime (hr)", ylab = "density",
 main = "Weibull density using MLEs from the lifetime data")
rug(xmp04.30$lifetime, col = "blue")

```

Or:
```{r}
weibull <- function(x) dweibull(x, shape = 2.15, scale = 1289.34)

plot(weibull, 
     xlim = c(0, 3000), 
     col = "red", xlab = "lifetime (hr)", 
     ylab = "density",
     main = "Weibull density using MLEs from the lifetime data")

rug(xmp04.31$lifetime, col = "blue")

```

## Contouring the log-likelihood function

```{r}
grid <- matrix(0.0, nrow = 101, ncol = 101)
scvals <- seq(0.5, 3.5, len = 101) # scale parameter
shvals <- seq(500, 2500, len = 101) # shape parameter

for (i in seq(along = scvals)) {
  for (j in seq(along = shvals)) {
    grid[i,j] <- llfunc(c(scvals[i], shvals[j]))
  }
}

contour(scvals, shvals, grid, levels = 77:85)
points(mle$estimate[1], mle$estimate[2], pch = "+", cex = 1.5)
title(xlab = expression(alpha), ylab = expression(beta))
```

Or use levels calculated from the chi-square distribution
```{r}
# Or use levels calculated from the chi-square distribution
contour(scvals, shvals, grid,
 levels = mle$min + qchisq(c(0.5,0.8,0.9,0.95,0.99), 2),
 labels = paste(c(50,80,90,95,99), "%", sep = ""))
```

## Lessons from the Weibull example

* The likelihood function is the same as the probability density but with the parameters varying and the data ﬁxed.

* For a random sample, the log-likelihood is 
  `sum(d<distname>(<data>, par1, par2, ..., log = TRUE))`

* We minimize the negative of the log-likelihood
```
llfunc <- function(x)
      -sum(d<distname>(<data>, par1 = x[1], ..., log = TRUE))
mle <- nlm(llfunc, <starting estimates>, hessian = TRUE)
```
* The inverse of the hessian provides an estimate of the variance-covariance matrix.

* For two-parameter models we can evaluate a grid of log-likelihood values and get contours.

* Standard errors from the inverse hessian are not always realistic indications of the variability in the parameter estimates.


```{r}
weibull <- function(x) dweibull(x, shape = 2, scale = 1)

plot(weibull, 
     xlim = c(0, 2), 
     col = "red", xlab = "lifetime (hr)", 
     ylab = "density",
     main = "Weibull density")



```

```{r}
library(ggplot2)

funcs <- list(function(x) dweibull(x, shape = 1,   scale = 1),
              function(x) dweibull(x, shape = 2,   scale = 1),
              function(x) dweibull(x, shape = 3.5, scale = 1))

cols <- heat.colors(3, 1)

p <-ggplot() + 
  xlim(c(0, 2))  

for(i in 1:length(funcs))
    p <- p + stat_function(aes(y=0), 
                           fun = funcs[[i]], 
                           colour = cols[i])
p <- p + labs(x = "x", y = "f(x)")
print(p)
```




