---
title: "R Notebook"
output: html_notebook
---

## An R sesssion on the Weibull example
```{r}
library(Devore7)

data(xmp04.31)
str(xmp04.31)
```

The Weibull probability plot is not available directly in R. However, the plot can be created using the formula ln[− ln(1 − p)] for the 5th, 15th, . . . , and 95th percentiles as given in text. The sequence 0.05, 0.15, . . . , 0.95 is generated with seq(0.05, 0.95, 0.1).

```{r}
with(xmp04.31, 
     plot(log(-log(1 - seq(0.05, 0.95, 0.1))), log(lifetime)))
```

```{r}
# Reasonable starting estimates are shape = 1, scale = 1000
# Do a simple evaluation at this set of parameters￿
sum(dweibull(xmp04.31$lifetime, shape=1, scale=1000, log=TRUE))
```

```{r}
# Optimization functions minimize so use negative log-likelihood
llfunc <- function(x) { # express as a function
 -sum(dweibull(xmp04.31$lifetime, shape=x[1], scale=x[2], log=TRUE))
}

# apply non-linear minimization
mle <- nlm(llfunc, c(shape = 1, scale = 1000), hessian = TRUE)

```

## Results of the Weibull Example

```{r}
str(mle)
```


```{r}
solve(mle$hessian) # approximate variance-covariance matrix￿İ
```

We see that the maximum of the log-likelihood is -77.1 achieved
at $\alpha = 2.15$ and $\beta = 1289.34$. The approximate standard errors of the estimates are $0.5444 = \sqrt 0.29596$ and $199.6 = \sqrt 38935.82$. We can use the standard errors to determine a grid of $(\alpha, \beta)$ values for contouring the log-likelihood function.






